---
title: "Deforestation and Forest Conversion"
author: "Roudranil Das"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---

```{r}
library(tidyverse)
library(tidytext)
library(scales)
library(fuzzyjoin)
library(maps)
library(countrycode)
library(ggthemes)
```
 
We will be using the following dataset from TidyTuesday with R: deforestation.
[Link](https://github.com/rfordatascience/tidytuesday/tree/master/data/2021/2021-04-06)

Loading in the datasets:
```{r}
forest <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-04-06/forest.csv')
forest_area <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-04-06/forest_area.csv')
brazil_loss <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-04-06/brazil_loss.csv')
# soybean_use <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-04-06/soybean_use.csv')
# vegetable_oil <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-04-06/vegetable_oil.csv')
```

Our analysis consists of the following 5 datasets:

The forest dataset shows the change every 5 years in forest conservation in hectares by country. To deal with the data better, we convert the hectares to $km^2$. And we also select those entities with a valid country code.
```{r}
forest <- forest %>% 
    mutate(net_forest_conversion = net_forest_conversion / 100) %>% 
    filter(str_length(code) == 3) %>% 
    rename(country = entity)
```

1. maximum forest conversion by country (absolute values)
```{r}
forest %>%
    group_by(year) %>%
    slice_max(abs(net_forest_conversion), n = 10) %>%
    ungroup() %>%
    mutate(country = reorder_within(country, net_forest_conversion, year)) %>%
    ggplot(aes(net_forest_conversion, country,
               fill = net_forest_conversion > 0)) +
    geom_col() +
    facet_wrap(~ year, scales = "free_y") +
    scale_x_continuous(label = comma) +
    scale_y_reordered() +
    theme(legend.position = "none") +
    labs(title = "Where have we seen the greatest change in net forest area in the years?",
         x = "Net change in forest area over the years (in sq km)",
         y = "")

# ggsave(filename = "net_change_in_forest_area_1.png",
#        device = "png",
#        dpi = 300,
#        path = "doc/figures/")
```


```{r echo=FALSE, results='hide'}
# 2. maximum forest increase (not included in report)

# forest %>% 
#     group_by(year) %>% 
#     slice_max(net_forest_conversion, n = 5) %>% 
#     ungroup() %>% 
#     mutate(country = reorder_within(country, net_forest_conversion, year)) %>% 
#     ggplot(aes(net_forest_conversion, country, fill = country)) + 
#     geom_col() +
#     facet_wrap(~year, scales = "free_y") +
#     scale_x_continuous(label = comma) +
#     scale_y_reordered() +
#     theme(legend.position = "none") + 
#     labs(title = "Where did we see the most increase in forest every year?",
#          x = "Net increase in forest area over the years (in sq km)",
#          y = "")

# ggsave(filename = "net_increase_in_forest_area_2.png",
#        device = "png",
#        dpi = 300,
#        path = "doc/figures/")
```

3. maximum deforestation
```{r}
forest %>% 
    group_by(year) %>% 
    slice_min(net_forest_conversion, n = 5) %>% 
    ungroup() %>% 
    mutate(country = reorder_within(country, -net_forest_conversion, year)) %>% 
    ggplot(aes(net_forest_conversion, country, fill = country)) + 
    geom_col() +
    facet_wrap(~year, scales = "free_y") +
    scale_x_continuous(label = comma) +
    scale_y_reordered() +
    theme(legend.position = "none") + 
    labs(title = "Where did we see the most deforestation every year?",
         x = "Net increase in forest area over the years (in sq km)",
         y = "")
# ggsave(filename = "most_deforestation_3.png",
#        device = "png",
#        dpi = 300,
#        path = "doc/figures/")
```



```{r echo=FALSE, results='hide'}
# 4. maximum forest conversion by year (not included in report)

# forest %>% 
#     group_by(year) %>% 
#     summarise(total_forest_conversion = sum(net_forest_conversion)) %>% 
#     ggplot(aes(x = as.factor(year), weight = total_forest_conversion)) +
#     geom_bar(aes(fill = as.factor(year))) + 
#     theme(legend.position = "none") +
#     labs(title = "Total forest area conversion by year",
#          x = "Years",
#          y = "Net forest area converted (in sq km)")
```

5. 10 countries with max change in forest area
```{r}
forest_area <- forest_area %>% 
    filter(!year %in% c(1990, 1991, 1992)) %>% 
    filter(str_length(code) == 3) %>% 
    rename(country = entity)

forest_area %>% 
    filter(year %in% c(min(year), max(year))) %>% 
    pivot_wider(names_from = year, values_from = forest_area, names_prefix = "year_") %>% 
    mutate(change_in_forest_area = year_2020 - year_1993) %>% 
    select(country, code, change_in_forest_area) %>% 
    filter(!is.na(change_in_forest_area)) %>% 
    arrange(desc(abs(change_in_forest_area))) %>% 
    head(10) %>% 
    mutate(country = fct_reorder(country, -change_in_forest_area)) %>% 
    mutate(change_in_forest_area = change_in_forest_area/100) %>% 
    ggplot(aes(x = as.factor(country), weight = change_in_forest_area)) +
    geom_bar(aes(fill = change_in_forest_area > 0)) +
    theme(legend.position = "none",
          axis.text.x = element_text(angle = 45, vjust = 0.5)) +
    scale_y_continuous(labels = percent) +
    labs(x = "Countries",
         y = "Change in forest area (in %)")

# ggsave(filename = "doc/figures/max_change_forest_area_5.png",
#       device = "png",
#       dpi = 300)
```

```{r}
#countries_with_highest_average_area <- 
    
forest_area %>% 
    mutate(forest_area = forest_area/100) %>% 
    group_by(code) %>% 
    summarise(avg_forest_area = mean(forest_area)) %>% 
    arrange(desc(avg_forest_area)) %>% 
    head(15)
    
# forest_area %>% 
#     filter(code %in% c("RUS", "BRA", "CAN", "USA", "AUS", "IND", "PER", "MEX", "IDN")) %>% 
#     mutate(country = fct_reorder(country, -forest_area),
#            forest_area = forest_area/100) %>% 
#     ggplot(aes(x = year, y = forest_area, color = country)) +
#     geom_line(size=1) +
#     scale_y_continuous(labels = percent)

forest_area %>% 
    mutate(forest_area = forest_area/100) %>% 
    mutate(country = fct_lump(country, 9, w = forest_area)) %>% 
    group_by(country, year) %>% 
    summarise(forest_area = sum(forest_area), .groups = "drop") %>% 
    mutate(country = fct_reorder(country, -forest_area)) %>% 
    ggplot(aes(year, forest_area, fill = country)) +
    geom_area() + 
    scale_y_continuous(labels = percent) +
    expand_limits(y = 0) +
    labs(x = "Year",
         y = "Percentage of global forest area",
         fill = "Countries")

# ggsave(filename = "pct_global_forest_area_6.png",
#        device = "png",
#        dpi = 300,
#        path = "doc/figures/")
```

```{r}
brazil_loss <- brazil_loss %>%
    select(-entity, -code)

```

We see that max land lost is all due to pasture.
```{r}
brazil_loss %>%
    pivot_longer(cols = commercial_crops:small_scale_clearing, names_to = "cause", values_to = "lost_forest_area") %>% 
    group_by(year) %>% 
    slice_max(order_by = lost_forest_area) %>% 
    ungroup()

brazil_loss %>%
    pivot_longer(cols = commercial_crops:small_scale_clearing, names_to = "cause", values_to = "lost_forest_area") %>% 
    mutate(lost_forest_area = lost_forest_area/100) %>% 
    filter(!cause == "pasture") %>% 
    group_by(year) %>% 
    slice_max(order_by = lost_forest_area) %>% 
    ungroup() %>% 
    ggplot(aes(year, lost_forest_area, fill = cause)) +
    geom_bar(width = 1, stat = "identity", color = "white") +
    scale_y_continuous(aes(breaks = 0:nlevels(year), labels = cause)) +
    guides(color = "none") +
    labs(fill = "Cause of loss",
         x = "Year",
         y = "Forest area lost (in sq km)") +
    scale_y_continuous(labels = comma)

# ggsave(filename = "scnd_highest_loss_of_forest_cover_7.png",
#        device = "png",
#        dpi = 300,
#        path = "doc/figures/")
```


```{r echo=FALSE, results='hide'}
# line graph (not included in report)

# brazil_loss %>%
#     pivot_longer(cols = commercial_crops:small_scale_clearing, names_to = "cause", values_to = "lost_forest_area") %>% 
#     filter(!cause == "pasture") %>% 
#     mutate(lost_forest_area = lost_forest_area/100) %>% 
#     ggplot(aes(year, lost_forest_area, color = cause)) +
#     geom_line() 
    
```

Forest conversion map
```{r}
# maps::iso3166

country_data <- forest %>% 
    filter(year == 2010) %>% 
    inner_join(maps::iso3166, by = c(code = "a3"))

map_data("world") %>% 
    as_tibble() %>% 
    filter(region != "Antarctica") %>% 
    regex_left_join(country_data, by = c(region = "mapname")) %>% 
    ggplot(aes(long, lat, group = group , fill = net_forest_conversion)) +
    geom_polygon(color = "black", size = 0.1) +
    scale_fill_gradient2(low = "red", high = "green") +
    theme_map() +
    labs(fill = "Net forest conversion \n(in sq km)",
         x = "",
         y = "") +
    theme(axis.text.x = element_blank(),
          axis.text.y = element_blank())

# ggsave(filename = "world_map_conv_8.png",
#        device = "png",
#        dpi = 300,
#        path = "doc/figures/")
```
